### 2016-05-04 p03-ch01-2015-monitoring/sec99-appendix.tex
Fix ref/label for Chapter 04

\appendix
\label{appendix}

@@

\subsection{Operational Semantics} % (fold)
\label{sub:operational_semantics}
We use operational semantics to present the semantics of the runtime monitoring. 
The global configuration is
$$
\mathcal{C} = \langle \mathcal{S}, \mathcal{A}, \mathcal{M}, P \rangle
$$
which respectively composes of the services, agreements, monitors, and platform.
Note that there is only one platform in this configuration.
For any actor in the system, we use a local configuration as $C = \langle Q, \sigma \rangle$ in which $Q$ is the queue of the messages of the actor and $\sigma$ is a local state that the actor may use for its responsibilities.

@@

\subsection{A Timed Automata for Budget Compliance}
\label{sec:prop:budget}

Similar to Section \ref{sec:service:avail:fsm}, we use a timed automata (see Definition \ref{def:timed:automata}) to model the behavior of service budget compliance $\beta(s,\tau)$ (see Definition \ref{def:budget:compliance}) represented in Figure \ref{fig:fsm:budgetcomp}.
% 
\begin{figure}[h]
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=4cm,thick]
\node[state,accepting]        (A)   {$q_B$};
\node[state]  [right of=A]    (N)   {$q_{\overline{B}}$};
\path
(A) edge    [loop left, left]  node    {$\delta_M^1=(\bot,g_B,\tau)$}     (A)
(A) edge    [bend right, below]       node    {$\delta_M^2=(\dot{a},g_{\overline{B}},\tau)$}     (N)
(N) edge    [loop right, right]  node    {$\delta_M^3=(\dot{a},g_{\overline{B}},\tau)$}     (N)
(N) edge    [bend right, above]       node    {$\delta_M^4=(\bot,g_B,\tau)$}     (A)
;
\end{tikzpicture}
\caption{A timed automata for Budget Compliance $\beta(s,\tau)$}
\label{fig:fsm:budgetcomp}
\end{figure}
%

@@

\section{Distributed Service Sustainability} % (fold)
\label{sec:distributed_service_sustainability}
\begin{defn}[Service Sustainability $\gamma(s,\tau)$]
\label{def:service:sustain}
We define the service sustainability as:
\[
\gamma(s,\tau) = \alpha(s,\tau,t_c) \times \beta(s,\tau)
\]
that is the composition of service availability $\alpha(s,\tau,t_c)$ and budget compliance $\beta(s,\tau)$ over the time span $\tau$.
Note that under a fixed agreement $a$ for $s$, $\alpha(s,t)$ and $\beta(s,t)$ have an inverse relation to one another:
\[
\alpha(s,t) \uparrow \iff \beta(s,t) \downarrow
\]
because consuming resources for service $s$ for $\alpha$ mean more cost for $s$ under the same agreement $a$ that leads to lower $\beta$ (and vice versa).
\end{defn}

@@

\begin{theorem}[Budget Compliance Convergence]
\label{thm:budget:comp:conv}
The automata $M_\beta$ on budget compliance $\beta(s,\tau)$ makes a transition to state $q_B$ in a finite amount of time under Assumptions \ref{def:single:resource:type} and \ref{def:resource:cost:constant} given that $\tau \geq \hat{t}$ and agreement $a$ for service $s$ is fixed and not changing through time.
\end{theorem}

\begin{proof}[Budget Compliance Convergence]
Proof can be completed using same technique in Theorem \ref{thm:service:avail:conv}.
\newline$\Box$
\end{proof}

\subsection{A Timed Automata for Service Sustainability}
\label{sec:sustain}

@@

\begin{theorem}[Sustainable Service Convergence]
\label{thm:sustain:conv}
The sustainability of service $s$ converges to:
$$
\lim_{T \to \infty} \; \gamma(s, \tau) = (E(s,\tau) - \varepsilon) \times (E_B(s,\tau) - \varepsilon_{\euro})
$$
if and only if, agreement $a$ for service $s$ is \emph{modifiable} from a monitoring window $\tau$ to the next given that either $\alpha(s,\tau,t_c)$ or $\beta(s,\tau)$ fails in one $\tau$.
\end{theorem}

@@

Let $M_\alpha$ be in $q_A$ and $M_\beta$ in $q_B$ at a given time $T$.
At time $T$, if $g_{\overline{A}}$ holds, $M_\alpha$ makes a transition to $q_{\overline{A}}$ and starts to allocate a new resource $r$ for $s$.
At a later time $T' \geq T + \tau$, $M_\alpha$ makes a transition to $q_A$ based Theorem \ref{thm:service:avail:conv}.
However, if $g_{\overline{B}}$ holds at time $T'$, $M_\beta$ makes a transition to $q_{\overline{B}}$ because the newly allocated resource $r$ makes budget compliance $\beta(s,\tau)$ violate.
Based on Theorem \ref{thm:budget:comp:conv}, $M_\beta$ makes a transition to $q_B$ after deallocating at least one resource from $s$ at a later time $T'' \geq T' + \tau$.
At time $T''$, the system recurs the behavior at time $T$ and thus $\gamma(s,\tau)$ for every subsequent $\tau$ cannot converge to $(1-\varepsilon)\times(1-\varepsilon_{\euro})$.
Similarly, it can be reasoned how changes in $M_\beta$ lead to changes in $M_\alpha$ and eventually destabilize $M_\gamma$.

@@

\section{Application of Actor-based Monitoring Model}
\label{sec:app}

@@

Theorem \ref{thm:sustain:conv} proved that $\gamma(s,\tau)$ converges to $(E(s,\tau)-\varepsilon)\times(E_B(s,\tau)-\varepsilon_{\euro})$ given certain conditions.
The values of $E(s,\tau)$ and $E_B(s,\tau)$ are bound to $1$.
To simplify, we define $f(\varepsilon,\varepsilon_{\euro}) = (1 - \varepsilon)\times(1-\varepsilon_{\euro})$.
On its own, this function is interesting.
Figure \ref{fig:tolerance:func} presents the landscape of this function with $\varepsilon$ and $\varepsilon_{\euro}$ as its variables.
%

The following summarizes the characteristics of the function and how it affects $\gamma(s,\tau)$:
\begin{itemize}
\item The following property holds:
\[
(1-\varepsilon) \times (1-\varepsilon_{\euro}) = 1 \iff \varepsilon = 0 \;\lor\; \varepsilon_{\euro} = 0
\]
This property shows that choosing $0$ value for $\varepsilon$ or $\varepsilon_{\euro}$ makes $\gamma(s,\tau)$ more strict and less tolerant. 
Unless it is not a hard business requirement, the tolerance parameters must allow for room of error for either $\alpha(s,\tau,t_c)$ or $\beta(s,\tau)$; otherwise, the sustainability $\gamma(s,\tau)$ exposes more fragility.
\item Similarly, choosing $\varepsilon = 1$ or $\varepsilon_{\euro} = 1$ means that respectively $\alpha(s,\tau,t_c)$ and $\beta(s,\tau)$ is \emph{not} a business driver.
In this setting, the service sustainability $\gamma(s,\tau)$ depends on a single property (either $\alpha(s,\tau,t_c)$ or $\beta(s,\tau)$) which makes its maintenance easier.
Additionally, choosing such model allows to design a business in which exclusively either service availability or budget compliance is crucial but not both.
\item Apart from the corner values, the value of the function is \emph{confined} by the arcs that are projected on the ceiling of the box in Figure \ref{fig:tolerance:func}.
This means that the function values changes co-dependently based on the changes of $\varepsilon$ and $\varepsilon_{\euro}$; i.e. optimization of $\gamma(s,\tau)$ is freely achievable by changing each variable.
\item It should be also noted that the values for $\varepsilon$ and $\varepsilon_{\euro}$ should \emph{not} be so relaxed; 
i.e. values higher than $0.1$ would create a large range of error which may not be desired.
\end{itemize}

@@

The parameters $E(s,\tau)$ and $E_B(s,\tau)$ are significantly driven from a business perspective.
Such parameters enforce business-level constraints on the services that are also bound from a legal perspective.
In the past decade, there is a considerable research on how to specify and model such business constraints generally referred to as service level agreements (SLA).
A number of languages, frameworks and models have been developed to be able to model and specify SLAs for the deployment of services.
Among others, we refer to SLA$\star$~\cite{kearney2010sla}, WSLA~\cite{keller2003wsla}, and SLAng~\cite{lamanna2003slang}.
The focus of this research is not enter this domain and yet build on top it a model to be able to reason about ``distributed service properties'' such as service availability or budget compliance. 
In the following, we present a set of pseudo algorithms for the model in Section \ref{sec:modelling} that present how the monitoring can be generated based on the agreements that are specified. 
In Algorithm \ref{alg:gen:monitor}, the monitoring actor method \jtt{check} is presented that collects the measurements and publishes them to the platform.

@@

The top surface (including the arcs) is the 2-D projected values for $\Gamma$.
% 
\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
  view={45}{45},
  colorbar,
  3d box=complete,
  enlargelimits=false,
  domain=0:1,
  xlabel=$\varepsilon_\alpha$,
  ylabel=$\varepsilon_\beta$,
  zlabel=$\Gamma$
]
\addplot3[surf]
  {(1-x) * (1-y)};
\addplot3[contour gnuplot={contour dir=z, labels=false, draw color=black}, thick, z filter/.code=\def\pgfmathresult{1}]
  {(1-x) * (1-y)};
\end{axis}
\end{tikzpicture}
\caption{
The landscape for $\Gamma$.
} 
\label{fig:sustain:factor}
\end{figure}
%

@@

$\mathcal{M}$ can be partitioned in two \emph{disjoint} subsets, $\mathcal{M}^+$ and $\mathcal{M}^-$, such that $\mathcal{M}^+$ only containts those that use platform $P$ API to \jtt{allocate} a new resource to service $s$ whereas $\mathcal{M}^-$ only contains those that use platform $P$ API to \jtt{deallocate} a resource from service $s$:
\begin{itemize}
  \item $\forall \mathcal{M}_k \in \mathcal{M}^+$ Thoerem \ref{thm:service:avail:conv} holds for $\mathcal{M}_k$.
  \item $\forall \mathcal{M}_k \in \mathcal{M}^-$ Theorem \ref{thm:service:budget:conv} holds for $\mathcal{M}_k$.
  \item $\forall \mathcal{M}_k \in \mathcal{M}^+ \; , \mathcal{M}_l \in \mathcal{M}^- \;, \mathcal{M}_m = \mathcal{M}_k \times \mathcal{M}_l$ Theorem \ref{thm:service:sust} holds for $\mathcal{M}_m$. 
\end{itemize}

\subsection{A Timed Automata for Platform} 
\label{sec:automata:platform}

To model the behavior of the platform $P$ with a timed automata, we use the notion of ``deadline'' from Real-Time ABS~\cite{JaghooriBCS09,bjork2013:rtabs}.
A message is expected to complete (either successfully or with failure) within a specific amount of time, the deadline.
We choose monitoring window $\tau$ as the deadline for any message that is internally processed by platform $P$.
In practice, a deadline for a message is part of the implementation details of any platform API for an IaaS provider.
Figure \ref{fig:fsm:platform} presents a timed automata for platform $P$.

In the timed automata that we design for different elements of the system, the following apply.
Every timed automaton has only one state which is the start and end state.
We use $\dot{c}$ to denote the clock of the timed automaton.
The timed automaton makes a transition at a random time in every monitoring window $\tau$.
A transition may have two forms: \emph{periodic} by the clock or \emph{reactive} through an external event or API call.
For example, $M_P$ has a trasition for \jtt{allocate} a new resource which is reactive to an external platform API call.
And, it has a transition for updating service state $\sigma(s)$ which is periodic and regulated by the clock. 
Every reactive transition has a fixed deadline equal to $\tau$.
A reactive transition is presented with a ``dotted'' line in Figure~\ref{fig:fsm} whereas a periodic transition is presented with a ``solid'' line.
Every transition is presented as $\frac{\textsf{condition}}{\textsf{action}}$.
The \textsf{condition} is a boolean expression which evaluates to either \jtt{true} or \jtt{false}.
The \textsf{action} is an atomic operation or an asynchronous message using Real-Time ABS with a fixed deadline equal to $\tau$.
For every timed automaton, there is an implicit transition as: $\frac{\dot{c} \geq \tau}{\dot{c} \gets 0}$.

@@

% M_P
\subcaptionbox{{\small{$M_P$: Timed Automaton for Platform $P$}}\label{fig:fsm:platform}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]   
    (s)   {{\small$\dot{c} < \tau$}};
% \node[state]   [right of=s]      (e)   {$\dot{c} = t$};
\path[every loop/.style={looseness=7}]
(s) 
    edge    
    [loop left,style=densely dotted]  
    node {$\jtt{allocate}(s,n) \; \bfjtt{deadline}(\tau)$}          
    (s)
(s) 
    edge    
    [loop right,style=densely dotted]   
    node {$\jtt{deallocate}(s) \; \bfjtt{deadline}(\tau)$} 
    (s)
(s)
    edge
    [loop,out=150,in=120,looseness=10]
    node {$\sigma(s) \gets \jtt{getState}(s) \; \bfjtt{deadline}(\tau)$}
    (s)
(s)
    edge
    [loop,in=30,out=60,looseness=10]
    node  {$h(s) \gets \jtt{getHistory}(s) \; \bfjtt{deadline}(\tau)$}
    (s)
(s)
    edge
    [loop above]
    node {$\frac{\forall s\; \exists r: \; r \in h(s) \wedge r \not\in \sigma(s)}{\bfjtt{skip}}$}
    (s)
;
\end{tikzpicture}
}

% M_alpha
\subcaptionbox{{\small{$M_\alpha$: Timed Automaton $\alpha(s,\tau,t_c)$}}\label{fig:fsm:serviceavail}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]
(s)   
    {{\small$\dot{c} < \tau$}};
\path[every loop/.style={looseness=7}]
(s) 
    edge    
    [loop left]  
    node {$\frac{\alpha_\sigma(s,\tau) < 1 - \varepsilon_\alpha}{P \; ! \; \jtt{allocate}(s,n_r)}$}          
    (s)
(s) 
    edge    
    [loop right]   
    node {$\frac{1 - \varepsilon_\alpha \leq \alpha_\sigma(s,\tau) \leq 1 + \varepsilon_\alpha}{h(s) \gets \sigma(s)}$} 
    (s)
;
\end{tikzpicture}
}

% M_beta
\subcaptionbox{{\small{$M_\beta$: Timed Automaton for $\beta(s,\tau)$}}\label{fig:fsm:budgetcompliance}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]   (s)   {$\dot{c} < \tau$};
\path[every loop/.style={looseness=10}]
(s) 
    edge    
    [loop left]  
    node {$\frac{\beta(s,\tau) < 1 - \varepsilon_\beta}{P \; ! \; \jtt{deallocate}(s)}$}          
    (s)
(s) 
    edge    
    [loop right]   
    node {$\frac{1 - \varepsilon_\beta \leq \beta(s,\tau) \leq 1 + \varepsilon_\beta}{\sigma(s) \gets h(s)}$} 
    (s)
;
\end{tikzpicture}
}

% M_gamma
\subcaptionbox{{\small{$M_\gamma$: Timed Automaton for $\gamma(s,\tau)$}}\label{fig:fsm:sustain}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]   (s)   {$\dot{c} < \tau$};
\path[every loop/.style={looseness=10}]
(s) 
    edge    
    [loop left]  
    node {$\frac{\alpha_\sigma(s,\tau) < 1 - \varepsilon_\alpha}{P \; ! \; \jtt{allocate}(s,n_r)}$}          
    (s)
(s) 
    edge    
    [loop right]   
    node {$\frac{1 - \varepsilon_\alpha \leq \alpha_\sigma(s,\tau) \leq 1 + \varepsilon_\alpha}{h(s) \gets \sigma(s)}$} 
    (s)
(s) 
    edge    
    [loop,in=30,out=60,looseness=10]  
    node {$\frac{\beta(s,\tau) < 1 - \varepsilon_\beta}{P \; ! \; \jtt{deallocate}(s)}$}          
    (s)
(s) 
    edge    
    [loop,in=120,out=150,looseness=10]   
    node {$\frac{1 - \varepsilon_\beta \leq \beta(s,\tau) \leq 1 + \varepsilon_\beta}{\sigma(s) \gets h(s)}$} 
    (s)
;
\end{tikzpicture}
}

\caption{Timed automata for elements of the monitoring platform}
\label{fig:fsm}
\end{figure}

\subsection{A Timed Automata for Service Availability}
\label{sec:service:avail:fsm}

We use timed automata~\cite{alur:1994:timedautomata} $M_\alpha$ to manage service availability (see Definition \ref{def:service:availability}).
$M_\alpha$ is designed to specifically address an \emph{under-capacity} service:

@@

\subsection{A Timed Automata for Service Budget Compliance}
\label{sec:automata:budget}

Analogously, we define a timed automata $M_\beta$ for service budget compliance (Definition~\ref{def:budget:compliance}). The automaton is shown in Figure~\ref{fig:fsm:budgetcompliance}.  $M_\beta$ is specifically designed to address a service that is \emph{over-budget}.
%

@@

\begin{defn}[Service Budget Compliance Convergence]
\label{thm:service:budget:conv}
The timed automata $M_\beta$ for service budget compliance $\beta(s,\tau)$ ensures that service $s$ is budget compliant in a finite amount of time.
\end{defn}

>>>>

\appendix
\label{ch04:appendix}

@@

\subsection{Operational Semantics} % (fold)
\label{ch04:sub:operational_semantics}
We use operational semantics to present the semantics of the runtime monitoring. 
The global configuration is
$$
\mathcal{C} = \langle \mathcal{S}, \mathcal{A}, \mathcal{M}, P \rangle
$$
which respectively composes of the services, agreements, monitors, and platform.
Note that there is only one platform in this configuration.
For any actor in the system, we use a local configuration as $C = \langle Q, \sigma \rangle$ in which $Q$ is the queue of the messages of the actor and $\sigma$ is a local state that the actor may use for its responsibilities.

@@

\subsection{A Timed Automata for Budget Compliance}
\label{ch04:sec:prop:budget}

Similar to Section \ref{ch04:sec:service:avail:fsm}, we use a timed automata (see Definition \ref{ch04:def:timed:automata}) to model the behavior of service budget compliance $\beta(s,\tau)$ (see Definition \ref{ch04:def:budget:compliance}) represented in Figure \ref{ch04:fig:fsm:budgetcomp}.
% 
\begin{figure}[h]
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=4cm,thick]
\node[state,accepting]        (A)   {$q_B$};
\node[state]  [right of=A]    (N)   {$q_{\overline{B}}$};
\path
(A) edge    [loop left, left]  node    {$\delta_M^1=(\bot,g_B,\tau)$}     (A)
(A) edge    [bend right, below]       node    {$\delta_M^2=(\dot{a},g_{\overline{B}},\tau)$}     (N)
(N) edge    [loop right, right]  node    {$\delta_M^3=(\dot{a},g_{\overline{B}},\tau)$}     (N)
(N) edge    [bend right, above]       node    {$\delta_M^4=(\bot,g_B,\tau)$}     (A)
;
\end{tikzpicture}
\caption{A timed automata for Budget Compliance $\beta(s,\tau)$}
\label{ch04:fig:fsm:budgetcomp}
\end{figure}
%

@@

\section{Distributed Service Sustainability} % (fold)
\label{ch04:sec:distributed_service_sustainability}
\begin{defn}[Service Sustainability $\gamma(s,\tau)$]
\label{ch04:def:service:sustain}
We define the service sustainability as:
\[
\gamma(s,\tau) = \alpha(s,\tau,t_c) \times \beta(s,\tau)
\]
that is the composition of service availability $\alpha(s,\tau,t_c)$ and budget compliance $\beta(s,\tau)$ over the time span $\tau$.
Note that under a fixed agreement $a$ for $s$, $\alpha(s,t)$ and $\beta(s,t)$ have an inverse relation to one another:
\[
\alpha(s,t) \uparrow \iff \beta(s,t) \downarrow
\]
because consuming resources for service $s$ for $\alpha$ mean more cost for $s$ under the same agreement $a$ that leads to lower $\beta$ (and vice versa).
\end{defn}

@@

\begin{theorem}[Budget Compliance Convergence]
\label{ch04:thm:budget:comp:conv}
The automata $M_\beta$ on budget compliance $\beta(s,\tau)$ makes a transition to state $q_B$ in a finite amount of time under Assumptions \ref{ch04:def:single:resource:type} and \ref{ch04:def:resource:cost:constant} given that $\tau \geq \hat{t}$ and agreement $a$ for service $s$ is fixed and not changing through time.
\end{theorem}

\begin{proof}[Budget Compliance Convergence]
Proof can be completed using same technique in Theorem \ref{ch04:thm:service:avail:conv}.
\newline$\Box$
\end{proof}

\subsection{A Timed Automata for Service Sustainability}
\label{ch04:sec:sustain}

@@

\begin{theorem}[Sustainable Service Convergence]
\label{ch04:thm:sustain:conv}
The sustainability of service $s$ converges to:
$$
\lim_{T \to \infty} \; \gamma(s, \tau) = (E(s,\tau) - \varepsilon) \times (E_B(s,\tau) - \varepsilon_{\euro})
$$
if and only if, agreement $a$ for service $s$ is \emph{modifiable} from a monitoring window $\tau$ to the next given that either $\alpha(s,\tau,t_c)$ or $\beta(s,\tau)$ fails in one $\tau$.
\end{theorem}

@@

Let $M_\alpha$ be in $q_A$ and $M_\beta$ in $q_B$ at a given time $T$.
At time $T$, if $g_{\overline{A}}$ holds, $M_\alpha$ makes a transition to $q_{\overline{A}}$ and starts to allocate a new resource $r$ for $s$.
At a later time $T' \geq T + \tau$, $M_\alpha$ makes a transition to $q_A$ based Theorem \ref{ch04:thm:service:avail:conv}.
However, if $g_{\overline{B}}$ holds at time $T'$, $M_\beta$ makes a transition to $q_{\overline{B}}$ because the newly allocated resource $r$ makes budget compliance $\beta(s,\tau)$ violate.
Based on Theorem \ref{ch04:thm:budget:comp:conv}, $M_\beta$ makes a transition to $q_B$ after deallocating at least one resource from $s$ at a later time $T'' \geq T' + \tau$.
At time $T''$, the system recurs the behavior at time $T$ and thus $\gamma(s,\tau)$ for every subsequent $\tau$ cannot converge to $(1-\varepsilon)\times(1-\varepsilon_{\euro})$.
Similarly, it can be reasoned how changes in $M_\beta$ lead to changes in $M_\alpha$ and eventually destabilize $M_\gamma$.

@@

\section{Application of Actor-based Monitoring Model}
\label{ch04:sec:app}

@@

Theorem \ref{ch04:thm:sustain:conv} proved that $\gamma(s,\tau)$ converges to $(E(s,\tau)-\varepsilon)\times(E_B(s,\tau)-\varepsilon_{\euro})$ given certain conditions.
The values of $E(s,\tau)$ and $E_B(s,\tau)$ are bound to $1$.
To simplify, we define $f(\varepsilon,\varepsilon_{\euro}) = (1 - \varepsilon)\times(1-\varepsilon_{\euro})$.
On its own, this function is interesting.
Figure \ref{ch04:fig:tolerance:func} presents the landscape of this function with $\varepsilon$ and $\varepsilon_{\euro}$ as its variables.
%

The following summarizes the characteristics of the function and how it affects $\gamma(s,\tau)$:
\begin{itemize}
\item The following property holds:
\[
(1-\varepsilon) \times (1-\varepsilon_{\euro}) = 1 \iff \varepsilon = 0 \;\lor\; \varepsilon_{\euro} = 0
\]
This property shows that choosing $0$ value for $\varepsilon$ or $\varepsilon_{\euro}$ makes $\gamma(s,\tau)$ more strict and less tolerant. 
Unless it is not a hard business requirement, the tolerance parameters must allow for room of error for either $\alpha(s,\tau,t_c)$ or $\beta(s,\tau)$; otherwise, the sustainability $\gamma(s,\tau)$ exposes more fragility.
\item Similarly, choosing $\varepsilon = 1$ or $\varepsilon_{\euro} = 1$ means that respectively $\alpha(s,\tau,t_c)$ and $\beta(s,\tau)$ is \emph{not} a business driver.
In this setting, the service sustainability $\gamma(s,\tau)$ depends on a single property (either $\alpha(s,\tau,t_c)$ or $\beta(s,\tau)$) which makes its maintenance easier.
Additionally, choosing such model allows to design a business in which exclusively either service availability or budget compliance is crucial but not both.
\item Apart from the corner values, the value of the function is \emph{confined} by the arcs that are projected on the ceiling of the box in Figure \ref{ch04:fig:tolerance:func}.
This means that the function values changes co-dependently based on the changes of $\varepsilon$ and $\varepsilon_{\euro}$; i.e. optimization of $\gamma(s,\tau)$ is freely achievable by changing each variable.
\item It should be also noted that the values for $\varepsilon$ and $\varepsilon_{\euro}$ should \emph{not} be so relaxed; 
i.e. values higher than $0.1$ would create a large range of error which may not be desired.
\end{itemize}

@@

The parameters $E(s,\tau)$ and $E_B(s,\tau)$ are significantly driven from a business perspective.
Such parameters enforce business-level constraints on the services that are also bound from a legal perspective.
In the past decade, there is a considerable research on how to specify and model such business constraints generally referred to as service level agreements (SLA).
A number of languages, frameworks and models have been developed to be able to model and specify SLAs for the deployment of services.
Among others, we refer to SLA$\star$~\cite{kearney2010sla}, WSLA~\cite{keller2003wsla}, and SLAng~\cite{lamanna2003slang}.
The focus of this research is not enter this domain and yet build on top it a model to be able to reason about ``distributed service properties'' such as service availability or budget compliance. 
In the following, we present a set of pseudo algorithms for the model in Section \ref{ch04:sec:modelling} that present how the monitoring can be generated based on the agreements that are specified. 
In Algorithm \ref{ch04:alg:gen:monitor}, the monitoring actor method \jtt{check} is presented that collects the measurements and publishes them to the platform.

@@

The top surface (including the arcs) is the 2-D projected values for $\Gamma$.
% 
\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
  view={45}{45},
  colorbar,
  3d box=complete,
  enlargelimits=false,
  domain=0:1,
  xlabel=$\varepsilon_\alpha$,
  ylabel=$\varepsilon_\beta$,
  zlabel=$\Gamma$
]
\addplot3[surf]
  {(1-x) * (1-y)};
\addplot3[contour gnuplot={contour dir=z, labels=false, draw color=black}, thick, z filter/.code=\def\pgfmathresult{1}]
  {(1-x) * (1-y)};
\end{axis}
\end{tikzpicture}
\caption{
The landscape for $\Gamma$.
} 
\label{ch04:fig:sustain:factor}
\end{figure}
%

@@

$\mathcal{M}$ can be partitioned in two \emph{disjoint} subsets, $\mathcal{M}^+$ and $\mathcal{M}^-$, such that $\mathcal{M}^+$ only containts those that use platform $P$ API to \jtt{allocate} a new resource to service $s$ whereas $\mathcal{M}^-$ only contains those that use platform $P$ API to \jtt{deallocate} a resource from service $s$:
\begin{itemize}
  \item $\forall \mathcal{M}_k \in \mathcal{M}^+$ Thoerem \ref{ch04:thm:service:avail:conv} holds for $\mathcal{M}_k$.
  \item $\forall \mathcal{M}_k \in \mathcal{M}^-$ Theorem \ref{ch04:thm:service:budget:conv} holds for $\mathcal{M}_k$.
  \item $\forall \mathcal{M}_k \in \mathcal{M}^+ \; , \mathcal{M}_l \in \mathcal{M}^- \;, \mathcal{M}_m = \mathcal{M}_k \times \mathcal{M}_l$ Theorem \ref{ch04:thm:service:sust} holds for $\mathcal{M}_m$. 
\end{itemize}

\subsection{A Timed Automata for Platform} 
\label{ch04:sec:automata:platform}

To model the behavior of the platform $P$ with a timed automata, we use the notion of ``deadline'' from Real-Time ABS~\cite{JaghooriBCS09,bjork2013:rtabs}.
A message is expected to complete (either successfully or with failure) within a specific amount of time, the deadline.
We choose monitoring window $\tau$ as the deadline for any message that is internally processed by platform $P$.
In practice, a deadline for a message is part of the implementation details of any platform API for an IaaS provider.
Figure \ref{ch04:fig:fsm:platform} presents a timed automata for platform $P$.

In the timed automata that we design for different elements of the system, the following apply.
Every timed automaton has only one state which is the start and end state.
We use $\dot{c}$ to denote the clock of the timed automaton.
The timed automaton makes a transition at a random time in every monitoring window $\tau$.
A transition may have two forms: \emph{periodic} by the clock or \emph{reactive} through an external event or API call.
For example, $M_P$ has a trasition for \jtt{allocate} a new resource which is reactive to an external platform API call.
And, it has a transition for updating service state $\sigma(s)$ which is periodic and regulated by the clock. 
Every reactive transition has a fixed deadline equal to $\tau$.
A reactive transition is presented with a ``dotted'' line in Figure~\ref{ch04:fig:fsm} whereas a periodic transition is presented with a ``solid'' line.
Every transition is presented as $\frac{\textsf{condition}}{\textsf{action}}$.
The \textsf{condition} is a boolean expression which evaluates to either \jtt{true} or \jtt{false}.
The \textsf{action} is an atomic operation or an asynchronous message using Real-Time ABS with a fixed deadline equal to $\tau$.
For every timed automaton, there is an implicit transition as: $\frac{\dot{c} \geq \tau}{\dot{c} \gets 0}$.

@@

% M_P
\subcaptionbox{{\small{$M_P$: Timed Automaton for Platform $P$}}\label{ch04:fig:fsm:platform}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]   
    (s)   {{\small$\dot{c} < \tau$}};
% \node[state]   [right of=s]      (e)   {$\dot{c} = t$};
\path[every loop/.style={looseness=7}]
(s) 
    edge    
    [loop left,style=densely dotted]  
    node {$\jtt{allocate}(s,n) \; \bfjtt{deadline}(\tau)$}          
    (s)
(s) 
    edge    
    [loop right,style=densely dotted]   
    node {$\jtt{deallocate}(s) \; \bfjtt{deadline}(\tau)$} 
    (s)
(s)
    edge
    [loop,out=150,in=120,looseness=10]
    node {$\sigma(s) \gets \jtt{getState}(s) \; \bfjtt{deadline}(\tau)$}
    (s)
(s)
    edge
    [loop,in=30,out=60,looseness=10]
    node  {$h(s) \gets \jtt{getHistory}(s) \; \bfjtt{deadline}(\tau)$}
    (s)
(s)
    edge
    [loop above]
    node {$\frac{\forall s\; \exists r: \; r \in h(s) \wedge r \not\in \sigma(s)}{\bfjtt{skip}}$}
    (s)
;
\end{tikzpicture}
}

% M_alpha
\subcaptionbox{{\small{$M_\alpha$: Timed Automaton $\alpha(s,\tau,t_c)$}}\label{ch04:fig:fsm:serviceavail}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]
(s)   
    {{\small$\dot{c} < \tau$}};
\path[every loop/.style={looseness=7}]
(s) 
    edge    
    [loop left]  
    node {$\frac{\alpha_\sigma(s,\tau) < 1 - \varepsilon_\alpha}{P \; ! \; \jtt{allocate}(s,n_r)}$}          
    (s)
(s) 
    edge    
    [loop right]   
    node {$\frac{1 - \varepsilon_\alpha \leq \alpha_\sigma(s,\tau) \leq 1 + \varepsilon_\alpha}{h(s) \gets \sigma(s)}$} 
    (s)
;
\end{tikzpicture}
}

% M_beta
\subcaptionbox{{\small{$M_\beta$: Timed Automaton for $\beta(s,\tau)$}}\label{ch04:fig:fsm:budgetcompliance}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]   (s)   {$\dot{c} < \tau$};
\path[every loop/.style={looseness=10}]
(s) 
    edge    
    [loop left]  
    node {$\frac{\beta(s,\tau) < 1 - \varepsilon_\beta}{P \; ! \; \jtt{deallocate}(s)}$}          
    (s)
(s) 
    edge    
    [loop right]   
    node {$\frac{1 - \varepsilon_\beta \leq \beta(s,\tau) \leq 1 + \varepsilon_\beta}{\sigma(s) \gets h(s)}$} 
    (s)
;
\end{tikzpicture}
}

% M_gamma
\subcaptionbox{{\small{$M_\gamma$: Timed Automaton for $\gamma(s,\tau)$}}\label{ch04:fig:fsm:sustain}}
{
\begin{tikzpicture}[->,>=stealth',auto,node distance=6cm,thick,scale=0.7,every node/.style={scale=0.7}]
\node[state,accepting]   (s)   {$\dot{c} < \tau$};
\path[every loop/.style={looseness=10}]
(s) 
    edge    
    [loop left]  
    node {$\frac{\alpha_\sigma(s,\tau) < 1 - \varepsilon_\alpha}{P \; ! \; \jtt{allocate}(s,n_r)}$}          
    (s)
(s) 
    edge    
    [loop right]   
    node {$\frac{1 - \varepsilon_\alpha \leq \alpha_\sigma(s,\tau) \leq 1 + \varepsilon_\alpha}{h(s) \gets \sigma(s)}$} 
    (s)
(s) 
    edge    
    [loop,in=30,out=60,looseness=10]  
    node {$\frac{\beta(s,\tau) < 1 - \varepsilon_\beta}{P \; ! \; \jtt{deallocate}(s)}$}          
    (s)
(s) 
    edge    
    [loop,in=120,out=150,looseness=10]   
    node {$\frac{1 - \varepsilon_\beta \leq \beta(s,\tau) \leq 1 + \varepsilon_\beta}{\sigma(s) \gets h(s)}$} 
    (s)
;
\end{tikzpicture}
}

\caption{Timed automata for elements of the monitoring platform}
\label{ch04:fig:fsm}
\end{figure}

\subsection{A Timed Automata for Service Availability}
\label{ch04:sec:service:avail:fsm}

We use timed automata~\cite{alur:1994:timedautomata} $M_\alpha$ to manage service availability (see Definition \ref{ch04:def:service:availability}).
$M_\alpha$ is designed to specifically address an \emph{under-capacity} service:

@@

\subsection{A Timed Automata for Service Budget Compliance}
\label{ch04:sec:automata:budget}

Analogously, we define a timed automata $M_\beta$ for service budget compliance (Definition~\ref{ch04:def:budget:compliance}). The automaton is shown in Figure~\ref{ch04:fig:fsm:budgetcompliance}.  $M_\beta$ is specifically designed to address a service that is \emph{over-budget}.
%

@@

\begin{defn}[Service Budget Compliance Convergence]
\label{ch04:thm:service:budget:conv}
The timed automata $M_\beta$ for service budget compliance $\beta(s,\tau)$ ensures that service $s$ is budget compliant in a finite amount of time.
\end{defn}
